---
title: "Homework 6"
author: "Donald Kane"
date: "March 28, 2022"
output:
  github_document
---

```{r message=FALSE, warning=FALSE}
#install.packages("randomForest")
library(randomForest)
```

```{r}
train <- read.csv(url("https://web.stanford.edu/~hastie/ElemStatLearn/datasets/vowel.train"))
train <- train[,2:ncol(train)]
```


# Question 1
### Convert the response variable in the “vowel.train” data frame to a factor variable prior to training, so that “randomForest” does classification rather than regression.

```{r}
train$y <- as.factor(train$y)
```

# Question 2
### Review the documentation for the “randomForest” function.
# Question 3
### Fit the random forest model to the vowel data using all of the 11 features using the default values of the tuning parameters.

```{r}
fit <- randomForest(y ~ ., data=train)
print(fit)          ## summary of fit object
plot(fit)           ## plot OOB MSE as function of # of trees
importance(fit)     ## variable importance 
varImpPlot(fit)     ## variable importance plot
```

# Question 4
### Use 5-fold CV and tune the model by performing a grid search for the following tuning parameters: 1) the number of variables randomly sampled as candidates at each split; consider values 3, 4, and 5, and 2) the minimum size of terminal nodes; consider a sequence (1, 5, 10, 20, 40, and 80).

```{r}

```

# Question 5
### With the tuned model, make predictions using the majority vote method, and compute the misclassification rate using the ‘vowel.test’ data.

```{r}

```















