---
title: "Homework 5"
author: "Donald Kane"
date: "February 22, 2022"
output:
  github_document
---

```{r}
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'
```

``` {r}
y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)
```

# Question 1
### Randomly split the mcycle data into training (75%) and validation (25%) subsets.

```{r}
# create training and testing data
samp = sort(sample(nrow(mcycle), nrow(mcycle)*.75))
train_x <- matrix(x[samp,])
test_x <- matrix(x[-samp,])
train_y <- y[samp]
test_y <- y[-samp]
```

# Question 2
### Using the mcycle data, consider predicting the mean acceleration as a function of time. Use the Nadaraya-Watson method with the k-NN kernel function to create a series of prediction models by varying the tuning parameter over a sequence of values. (hint: the script already implements this)

```{r}
kernel_epanechnikov <- function(x, x0, lambda=1) {
  d <- function(t)
    ifelse(t <= 1, 3/4*(1-t^2), 0)
  z <- t(t(x) - x0)
  d(sqrt(rowSums(z*z))/lambda)
}

kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))
  w <- rep(0, length(d))
    w[order(d)[1:k]] <- 1
  
  return(w)
}

nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  attr(yhat, 'k') <- k
  return(yhat)
}

## create a grid of inputs 
x_plot <- matrix(seq(min(x),max(x),length.out=100),100,1)
y_hat_plot <- nadaraya_watson(y, x, x_plot,
  kernel_epanechnikov, lambda=1)

y_hat <- nadaraya_watson(y, x, x,
  kern=kernel_k_nearest_neighbors, k=10)

y_hat_plot <- nadaraya_watson(y, x, x_plot,
  kern=kernel_k_nearest_neighbors, k=10)

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_plot, col="green", lwd=2)
```

# Question 3
### With the squared-error loss function, compute and plot the training error, AIC, BIC, and validation error (using the validation data) as functions of the tuning parameter.

```{r}
effective_df <- function(y, x, kern, ...) {
  y_hat <- nadaraya_watson(y, x, x,
    kern=kern, ...)
  sum(diag(attr(y_hat, 'k')))
}

loss_squared_error <- function(y, yhat)
  (y - yhat)^2

error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d
```

``` {r}
y_hat <- nadaraya_watson(train_y, train_x, train_x,
  kern=kernel_k_nearest_neighbors, k=1)
y_hat_test <- nadaraya_watson(test_y, test_x, test_x,
  kern=kernel_k_nearest_neighbors, k=1)

eff_df <- effective_df(train_y, train_x, kern = kernel_k_nearest_neighbors, k = 1)

aic1 <- aic(train_y, y_hat, eff_df)
bic1 <- bic(train_y, y_hat, eff_df)
train_error <- error(train_y, y_hat)
val_error <- error(test_y, y_hat_test)

plot(train_x, train_y, xlab="Time (ms)", ylab="Acceleration (g)")
lines(x_plot, y_hat_plot, col="green", lwd=2)
legend('topleft', legend = c(paste0("aic = ", aic1), paste0("bic = ", bic1), paste0("Trainning Error = ", train_error), paste0("Validation Error = ", val_error)), bty = 'n', cex = 0.5)
```

# Question 4
### For each value of the tuning parameter, Perform 5-fold cross-validation using the combined training and validation data. This results in 5 estimates of test error per tuning parameter value.

```{r}
#install.packages("caret")
library(caret)
```

``` {r}
## 5-fold cross-validation of knnreg model
## create five folds
set.seed(1985)
mcycle_acc  <- createFolds(mcycle$accel, k=5)
print(mcycle_acc)
sapply(mcycle_acc, length)  ## not all the same length

cvknnreg <- function(kNN = 10, flds=mcycle_acc) {
  cverr <- rep(NA, length(flds))
  for(tst_idx in 1:length(flds)) { ## for each fold
    
    ## get training and testing data
    mcycle_trn <- mcycle[-flds[[tst_idx]],]
    mcycle_tst <- mcycle[ flds[[tst_idx]],]
    
    ## fit kNN model to training data
    knn_fit <- knnreg(accel ~ times,
                      k=kNN, data=mcycle_trn)
    
    ## compute test error on testing data
    pre_tst <- predict(knn_fit, mcycle_tst)
    cverr[tst_idx] <- mean((mcycle_tst$accel - pre_tst)^2)
  }
  return(cverr)
}

## Compute 5-fold CV for kNN = 1:20
cverrs <- sapply(1:20, cvknnreg)
print(cverrs) ## rows are k-folds (1:5), cols are kNN (1:20)
(cverrs_mean <- apply(cverrs, 2, mean))
(cverrs_sd   <- apply(cverrs, 2, sd))
```

# Question 5
### Plot the CV-estimated test error (average of the five estimates from each fold) as a function of the tuning parameter. Add vertical line segments to the figure (using the segments function in R) that represent one “standard error” of the CV-estimated test error (standard deviation of the five estimates from each fold).

```{r}
## Plot the results of 5-fold CV for kNN = 1:20
plot(x=1:20, y=cverrs_mean, 
     ylim=range(cverrs),
     xlab="'k' in kNN", ylab="CV Estimate of Test Error")
segments(x0=1:20, x1=1:20,
         y0=cverrs_mean-cverrs_sd,
         y1=cverrs_mean+cverrs_sd)
best_idx <- which.min(cverrs_mean)
points(x=best_idx, y=cverrs_mean[best_idx], pch=20)
abline(h=cverrs_mean[best_idx] + cverrs_sd[best_idx], lty=3)
```

# Question 6
### Interpret the resulting figures and select a suitable value for the tuning parameter.

From the graph above in question 4, the minimum test error is at k = 14. Using the one standard error rule, the suitable value for the tuning parameter would be k = 20 because as k increases, the model becomes less flexible.














